---
title:
output: html_document
runtime: shiny
---


##

<!-- Estimation Period -->


```{r, messsage=FALSE, warning=FALSE, echo = FALSE}
#{{{
## in the process of translating this code into a Shiny app
## just run for each pair?
library(dygraphs)
shinyUI(fluidPage(
    ## titlePanel("submitButton example"),
    fluidRow(
        column(3, wellPanel(
                      ## inputPanel( 
                      ## dateRangeInput("estim_period", "Estimation Period", start = "2015-09-01", end = "2015-12-31"),
                      dateInput("estim_period_start", "Estimation Period Start", value = "2015-08-01"), 
                      dateInput("estim_period_end", "End", value = "2016-01-01"), 
                      ## selectizeInput(tickers, "Enter Tickers", )
                      textInput("ticker1", "First Stock Ticker", value = "XOM"),
                      textInput("ticker2", "Second Stock Ticker", value = "CVX")
                  )),
        column(9,
               renderDygraph({
               library(quantmod)
               ## Estimation period
               ## from <- input$estim_period[1]
               ## to <- input$estim_period[2]

               from <- input$estim_period_start
               to <- input$estim_period_end

               ## The code below to calculate an xts object of returns is from an SO or Quant.SE post.  I need to find it and cite.
               env <- new.env()
               Symbols <- c(input$ticker1, input$ticker2)
               getSymbols(Symbols = Symbols, env = env, from = from, to = to)
               args <- eapply(env = env, FUN = function(x){ClCl(x)})[Symbols]
               returns <- na.omit(do.call(what = merge, args = args))
               colnames(returns) <- Symbols
               ## convert reutrns to  normalized prices
               norm.prices <<- cumsum(returns)

               norm.prices.df <<- as.data.frame(norm.prices)



                   update1 <- input$ticker1
                   update2 <- input$ticker2
                   update.e1 <- input$estim_period_start
                   update.e2 <- input$estim_period_end                   
                   dygraph(norm.prices) 
                                        #}}} 
               }),


               renderPrint({    
                   ## function to find the average squared deviation of the normalized prices for two stocks
                   update3 <- input$ticker1
                   update4 <- input$ticker2
                   update.e3 <- input$estim_period_start
                   update.e4 <- input$estim_period_end                   
                   
                   ## Calculate the estimation period parameters (the average squared deviation and the standard deviation of the squared deviations):
                   asd.e <<- mean(( norm.prices.df[,1] - norm.prices.df[,2])^2 )
                   ssd.e <<- sd(( norm.prices.df[,1] - norm.prices.df[,2])^2 )
                   cat("The average squared deviation is", asd.e ,"\n")
                   cat("The standard deviation of the squared deviations is:", ssd.e ,"\n")
                   ## paste("The average squared deviation is", asd.e ,"\n")
                   ## paste("The standard deviation of the squared deviations is:", ssd.e ,"\n")

               })
               )
    )
)
)
#}}}
```

## Trading Period

We monitor how many standard deviations the squared deviation is from its average squared deviation. 

-  When the squared deviation differs from its average by a threshold number of standard deviations, we'll initiate a pairs trade. 

-  We close the pairs trade when the squared deviation drops below another threshold number of standard deviations.

## Thresholds

Implementation of the trade requires thresholds at which you open and close your pairs trade. Gatev et al (2008) enter the trade at 2 standard deviations wide, and they close the trade when the prices cross.  

-  You can observe what others have done, but there is no theoretical reason why the threshold should be 2 instead of 1.9 or 2.1. You can, of course, run different thresholds over historical data and see what was optimal. 

-  The app below allows you to try different thresholds and see the profit.

-  Whatever thresholds were optimal may not be optimal over later trading periods.  

## Adjusting for Market Risk


If the two stocks are really good substitutes for one another, they likely have similar levels of market risk (measured by [beta coefficients](http://www.5minutefinance.org/concepts/the-capital-asset-pricing-model)). You can therefore buy and short equal dollar amounts. 

- However, if the stocks have substantially different beta coefficients (first, reconsider if it is a good pair), you may want to weight your investment in each so that the beta of the position is close to zero.

- For example, say stock A has a beta of 0.8 and stock B has a beta of 1.2. Then you want to invest $\frac{1.2}{0.8} = 1.5$ times as much in stock A as B. So, say you invest \$10,000 in stock B, then you want to invest \$15,000 in stock A.

Gatev et al (2008) do not make a market risk adjustment and so neither will the results below.  

##

```{r, messsage=FALSE, warning=FALSE, echo = FALSE}
#{{{
inputPanel(

    sliderInput("buy", "Enter Trade at Threshold:", min = .5, max = 4, value = 2, step = .1, animate = FALSE),

    sliderInput("sell", "Close Trade at Threshold", min = 0, max = 1.5, value = 0.5, step = .1, animate = FALSE)
)

renderDygraph({
from.t <- input$estim_period_end + 1
to.t <- Sys.Date()

## The code below to calculate an xts object of returns is from an SO or Quant.SE post.  I need to find it and cite.
env.t <- new.env()
Symbols <- c(input$ticker1, input$ticker2)
getSymbols(Symbols = Symbols, env = env.t, from = from.t, to = to.t)
args <- eapply(env = env.t, FUN = function(x){ClCl(x)})[Symbols]
returns.t <- na.omit(do.call(what = merge, args = args))
colnames(returns.t) <- Symbols
## convert reutrns to  normalized prices
norm.prices.t <<- cumsum(returns.t)## ll <- dim(norm.prices.df)[2]

numSD <- ( (norm.prices.t[,1] - norm.prices.t[,2])^2 - asd.e ) / ssd.e

## nope
## signal <- numSD
## signal[signal < 0] <- 1
## signal[signal > 0 & signal < 2] <- 0
## signal[signal > 2] <- 2
trade <- rep(0, length(seq_along(numSD)))
position <- rep(0, length(seq_along(numSD)))

for (i in 2:length(seq_along(numSD))) {
    if (numSD[i] > input$buy && numSD[i-1] < input$buy && position[i-1] < 1) {
        trade[i] <- 1
    } else {
        if (numSD[i] < input$sell && numSD[i-1] > input$sell && position[i-1] == 1) {
            trade[i] <- -1
        } else {
            trade[i] <- 0
        }
    }
    position[i] <- trade[i] + position[i-1]
}

### Now add profit calculation here ----
## First calculate how many standard deviations we made -----
stdev.when.trade <- trade * numSD
profitSD <- sum(stdev.when.trade)

if(position[length(position)] == 1) {
    profitSD = profitSD - numSD[length(numSD)]
} else {
    profitSD = profitSD
    }

## Now calculate the percent profit per standard deviation ----
### square root of the standard dev (now in units of deviations instead of sq devs) of squared devs times the number of standard devs
#### Don't know if this is the correct calculation -- probably not.
## amt.devs.percent <- 100 * sqrt(ssd.e) * profitSD

## This % profit calculation looks correct ----
## instead, just look at spread on that day in norm.prices.t
spread.when.trade <- trade * abs(norm.prices.t[,1] - norm.prices.t[,2])
profit.percent <- sum(spread.when.trade)


if(position[length(position)] == 1) {
    profit.percent = profit.percent - abs(norm.prices.t[dim(norm.prices.t)[1], 1] - norm.prices.t[dim(norm.prices.t)[1], 2])
} else {
    profit.percent = profit.percent
    }

profit.percent <- profit.percent * 100

data.dygraph <- cbind(numSD, as.xts(position, order.by = index(numSD)))
names(data.dygraph) <- c("# of Standard Deviations", "Position: 0 = no trade, 1 = trade on")

dygraph( data.dygraph  , main = paste0(colnames(returns.t)[1], " & ", colnames(returns.t)[2], ":  Profit in Standard Deviations: ", round(profitSD, 2) , ", Percent Profit ", round(profit.percent, 2), "%"), ylab = "# Standard Deviations") %>%
dyOptions(colors = RColorBrewer::brewer.pal(2, "Set1"))
#}}}    

})
```

##  Opening/Closing a Trade

Using the default pair, dates, and thresholds, you can see we open a pairs trade in the beginning of February (and close it in March).  How do we know which stock to buy and which to sell? 

-  In the next slide the normalized prices are charted over the trading period.  You can see when we open the trade, CVX's normalized price is below XOM's.  Therefore to open the trade we buy CVX and short an equal dollar amount of XOM.  

-  To close the pairs trade in March, we sell all of our CVX and buy back all of our XOM.

##

```{r, messsage=FALSE, warning=FALSE, echo = FALSE}
#{{{
library(dygraphs)
renderDygraph({
    update3 <- input$ticker1
    update4 <- input$ticker2
    update.e2 <- input$estim_period
    dygraph(norm.prices.t, main = "Normalized Prices Over the Trading Period") 
#}}}    
})
```
