---
title: "The Time Value of Money: Multiple Cash Flows"
date: "06/02/2015"
output: ioslides_presentation
runtime: shiny
smaller: true
---

<!-- author: "Matt Brigida: Milken Institute, Center for Financial Markets" -->
<!-- affiliation: "Milken Institute Center for Financial Markets" -->



## Introduction

In this presentation we'll cover some common methods of dealing with contracts that are comprised of multiple cash flows.


## Annuities

An **annuity** is a contract which pays a fixed amount at the end of each period for a fixed number of periods.  Many common financial contract are annuities, such as fixed rate mortgage and auto loans.  If the payments are in the beginning of a period (such as renting an apartment) the contract is known as an *annuity due*.

There are 
- $r$ the rate per period
- $n$ the number of periods
- $C$ the cash flow per period



## Future Value
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#{{{
inputPanel(
	   sliderInput("rate", label = "Annual Interest Rate", min = -.05, max =0.5, step = .01, value = .05),
	   sliderInput("years", label = "Number of Years", min = 1, max =50, step = 1, value = 10)
	   #            sliderInput("pv", label = "Present Value", min = 1, max =50, step = 1, value = 10),
	   )

renderPlot({
  fv <- (1 + input$rate)^{1:input$years}
  fv <- c(1, fv)
  plot(fv, type = 'l', lwd = 3, col = 3, ylab = "$", xlab = "Years")
  legend("topleft", "Future Value of $1", col = 3, lwd = 3, lty = 1)
})
#}}}
```


[Should we put a calculator which will return *FV*, *PV*, *r*, and *n* given the other inputs?]

<!---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#{{{
# inputPanel(
#            sliderInput("corr", label = "Stock Correlation", min = -1, max =1, step = 1, value = 1)
#            )
# 
# renderPlot({
#   par(mfrow = c(2,1), mai = c(.1,.8,.3,.1))
#   x <- seq(0, 2 * pi, by = .01)
#   plot(0.25 * sin(x), type = 'l', lwd = 2, col = 3, ylab = "Asset Return", ylim = c(-.3, .3), xlab = "", xaxt = 'n')
# legend("topright", c("Stock 1", "Stock 2"), col = c(3, 4), lwd = c(2,2), lty = c(1,2))
# 
#   y <- 0
#   if (input$corr == -1){
#     lines(-0.25 * sin(x), type = 'l', lwd = 2, col = 4, lty = 2)
#   } else {
#     if (input$corr == 0){
#       lines(0.25 * cos(x), type = 'l', lwd = 2, col = 4, lty = 2)
#     } else {
#       lines(0.25 * sin(x), type = 'l', lwd = 2, col = 4, lty = 2)
#     }
#   }
# 
#   if (input$corr == -1){
#     plot(0 * sin(x), ylim = c(-.3, .3), ylab = "Port. Return", type  = 'l', lwd = 2, xaxt = 'n', xlab = "Time")
#   } else {
#     if (input$corr == 0){
#       plot(0.5 * 0.25 * sin(x), ylim = c(-.3, .3), ylab = "Port. Return", type  = 'l', lwd = 2, xaxt = 'n', xlab = "Time")
#     } else {
#       plot(0.25 * sin(x), ylim = c(-.3, .3), ylab = "Port. Return", type  = 'l', lwd = 2, xaxt = 'n', xlab = "Time")
#     }
#   } 
#   })
#}}}
```
-->
